---
description: AFL Defense vs Position
title: Defense vs Position
---

```{css, echo=FALSE}
.box-score {
  font-family: 'Roboto', Helvetica, Arial, sans-serif;
}

.box-score-title {
  margin-top: 24px;
  padding: 8px;
  background-color: hsl(205, 100%, 36%);
  color: hsl(0, 0%, 98%);
  font-size: 15px;
  font-weight: 400;
}

```

```{r setup, include=FALSE}
library(tidyverse); library(reactable); library(shiny)

knitr::opts_chunk$set(echo = TRUE)
stats <- read.csv(file = '/Users/morts/Dropbox/Github-DFSAustralia-website/database/database-files/afl-database-2022.csv', stringsAsFactors = FALSE)

year_max <- max(stats$year)

make_color_pal <- function(colors, bias = 1) {
    get_color <- colorRamp(colors, bias = bias)
    function(x) rgb(get_color(x), maxColorValue = 255)
}

negative_color <- make_color_pal(c("#ffffff", "#9ae264"), bias = 2)
positive_color <- make_color_pal(c("#ffffff", "#ff6347"), bias = 2)

dvp_stats <- stats %>%
    filter(year == year_max) %>%
    filter(timeOnGroundPercentage >= 50) %>%
    group_by(player_id) %>%
    mutate(avgDT = mean(dreamTeamPoints)) %>%
    ungroup() %>%
    mutate(PM = dreamTeamPoints - avgDT) %>%
    group_by(position_group, opponent) %>%
    summarise(DVP = round(mean(PM),1)) %>%
    filter(!is.na(position_group))

max_dvp <- max(dvp_stats$DVP, na.rm = TRUE)
min_dvp <- min(dvp_stats$DVP, na.rm = TRUE)

dvp_stats_table <- dvp_stats %>%
    pivot_wider(names_from = position_group, values_from = DVP) %>%
    select(opponent, `Key Forward`, `Small/Medium Forward`, 
           `Inside Midfielder`, Ruck, `Wing/Attacking Defender`, `General Defender`, `Designated Kicker`)

# Define function for generating table

dvp_table <- reactable(
    dvp_stats_table,
    pagination = FALSE,
    fullWidth = TRUE,
    defaultSortOrder = "desc",
    wrap = FALSE,
    compact = TRUE,
    sortable = TRUE,
    striped = TRUE,
    defaultColDef = colDef(
        align = "center",
        minWidth = 80,
        style = function(value) {
            if (!is.numeric(value)) return()
            if(value < 0){
                list(background = positive_color(value/min_dvp), `font-weight` = "bold",
                     border =  "solid 1px", `border-color` = "white")
            } else if(value > 0){
                list(background = negative_color(value/max_dvp), `font-weight` = "bold",
                     border =  "solid 1px", `border-color` = "white")
            } else {list(`font-weight` = "bold", border =  "solid 1px", `border-color` = "white")}
        }
    ),
    columns = c(
        list(
            opponent = colDef(name = "Team", style = list(`font-weight` = "bold")),
            `Key Forward` = colDef(
                name = "KEY FWD",
                cell = function(value){
                    if(value == max(dvp_stats_table$`Key Forward`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Key Forward`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Key Forward`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            `Small/Medium Forward` = colDef(
                name = "GEN FWD",
                cell = function(value){
                    if(value == max(dvp_stats_table$`Small/Medium Forward`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Small/Medium Forward`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Small/Medium Forward`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            `Inside Midfielder` = colDef(
                name = "INS MID",
                cell = function(value){
                    if(value == max(dvp_stats_table$`Inside Midfielder`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Inside Midfielder`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Inside Midfielder`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            Ruck = colDef(
                name = "RUC",
                cell = function(value){
                    if(value == max(dvp_stats_table$Ruck)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$Ruck, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$Ruck, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            `Wing/Attacking Defender` = colDef(
                name = "WNG DEF",
                cell = function(value){
                    if(value == max(dvp_stats_table$`Wing/Attacking Defender`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Wing/Attacking Defender`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Wing/Attacking Defender`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            `General Defender` = colDef(
                name = "GEN DEF",
                cell = function(value){
                    if(value == max(dvp_stats_table$`General Defender`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`General Defender`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`General Defender`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            ),
            `Designated Kicker` = colDef(
                name = "DES KCK",
                cell = function(value){
                    if(value == max(dvp_stats_table$`Designated Kicker`)){
                        tagList(value, "  ", icon("trophy"), icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Designated Kicker`, TRUE)[2]){
                        tagList(value, "  ", icon("trophy"), icon("trophy"))
                    } else if(value == sort(dvp_stats_table$`Designated Kicker`, TRUE)[3]){
                        tagList(value, "  ", icon("trophy"))
                    } else {value}
                },
                html = TRUE
            )
        )
    )
)
  
```

**Defense vs Position (DVP)** is shown in terms of fantasy points conceded above (+) or below (-) a players average. For example, a DVP of +10.0 would indicate that that team allows an opposing player to score 10 points above their season average.

Positions in the DVP table have been classified by using a clustering algorithm similar to that discussed [here](https://www.dfsaustralia.com/blog/2020/08/23/2020-08-17-player-clustering-advanced-defence-v-position/){target="_blank"}. Refer to the glossary below for further explanation of each position.

If you find this useful consider **SUPPORTING US** by buying a coffee from as little as $5. This allows us to continue providing this great content for free.

<center>
<script type='text/javascript' src='https://storage.ko-fi.com/cdn/widget/Widget_2.js'></script><script type='text/javascript'>kofiwidget2.init('Buy a $5 Coffee', '#29abe0', 'E1E8C5INN');kofiwidget2.draw();</script> 
</center>

```{r, echo=FALSE}
  div(class = "box-score", style = "font-size:75%; white-space:nowrap",
      div(class = "box-score-title", tags$span("Defense vs Position (+/-)")),
      dvp_table
  )
```

\

**GLOSSARY**  
* **KEY FWD:** __Key Forward__ which is fairly explanatory (e.g. Harry McKay, Josh J. Kennedy, etc)  
* **GEN FWD:** __General Forward__ encompasses small/medium forwards aswell as half forwards that push up the ground from time to time (e.g. Cameron Zurhaar, Charlie Spargo, Matthew Owies, Ned McHenry, etc)  
* **INS MID:** __Inside Midfielders__ are your centre bounce midfielders (e.g. Brad Crouch, Christian Petracca, Clayton Oliver, Matthew Kennedy, etc)  
* **RUC:** __Ruck__ needs no further explanation  
* **WNG DEF:** includes both __Wing__ and __Attacking Defenders__ who push up the ground (e.g. Blake Acres, Daniel Howe, James Jordon, Tom Atkins, Dyson Heppell, Isaac Quaynor, etc)  
* **GEN DEF:** __General Defender__ includes key defenders and general defenders that aren't classified as Attacking Defenders above or Designated Kickers below (e.g. Darcy Gardiner, Griffin Logue, Luke Brown, Sam Taylor, etc)  
* **DES KCK:** __Designated Kicker__ refers to players who take the majority of their team's kick-ins after opposition goals (e.g. Bailey Dale, Alex Witherden, Jayden Short, Steven May, etc). These players could also be classified as GEN DEF or WNG DEF but have been given their own classification considering their fantasy scoring is affected by opposition shots on goal.

---

