---
title: AFLW Extended Stats
banner: img/banners/AFLW Logo.png
author: ''
date: '2022-01-10'
slug: aflw-extended-stats
categories: []
tags:
  - AFLW
---

```{r setup, include=FALSE}
library(tidyverse); library(reactable)
```

The AFLW season has kicked off for 2022 and people are starting to get interested in AFLW Fantasy. Moreira's Magic have released a great product that everyone should be getting behind [aflwfantasy.com.au](https://aflwfantasy.com.au/){target="_blank"}.

We don't envisage providing data tools at this stage but to do our little part in building this community we've collated the advanced stats for Round 1 of the AFLW season and made them downloadable [here](/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv).

The stats available include ruck contests attended, centre bounce attendances (CBAs), kickins taken and played on from, etc.

Head back over after the completion of each round and download updated data (if we don't get too lazy).

* [Download Round 1 Stats](/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv)
* [Download Round 2 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 2.csv) *(excludes BL v WB match stats as game has been [postponed](https://www.womens.afl/news/79797){target="_blank"})*
* [Download Round 3 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 3.csv) *(excludes GC v WB match stats as game has been [postponed](https://www.womens.afl/news/81199/r3-fixture-update-two-matches-cancelled-tuesday-night-clash-confirmed){target="_blank"})*
* [Download Round 4 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 4.csv) *(excludes STK v WCE match stats as game has been [postponed](https://www.womens.afl/news/82183/saints-eagles-clash-called-off-due-to-health-and-safety-protocols){target="_blank"})*
* [Download Round 4.5 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 4.5.csv) *(stats for WB v FRE only, game [brought forward](https://www.womens.afl/news/82357/what-s-in-a-round-footy-s-new-seven-day-schedule-explained){target="_blank"} from Round 10)*
* [Download Round 5 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 5.csv) *(excludes GWS v STK match stats as game has been [postponed](https://www.womens.afl/news/83204/giants-hit-by-health-and-safety-protocols-saints-game-postponed){target="_blank"})*
* [Download Round 5.5 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 5.5.csv) *(stats for STK v WCE only, [replacement](https://www.womens.afl/news/82183/saints-eagles-clash-called-off-due-to-health-and-safety-protocols){target="_blank"} game for Round 4)*
* [Download Round 6 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 6.csv)
* [Download Round 6.5 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 6.5.csv) *(stats for GC v BL only, game [brought forward](https://www.womens.afl/news/85624/lions-rack-up-highest-score-of-season-in-massive-win){target="_blank"} from Round 11)*
* [Download Round 7 Stats](/data/blog - AFLW Extended Stats Round 1/Extended AFLW Stats Round 7.csv)

Head over to [www.womens.afl](https://www.womens.afl/matches/2022){target="_blank"} for season fixtures.

***
#### Round 1 Extended Stats
```{r, echo=FALSE, message=FALSE}
stats <- read.csv("../../static/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv")
stats <- stats %>% filter(fantasy > 0)
reactable(stats,
          pagination = FALSE,
          compact = TRUE,
          highlight = TRUE,
          wrap = FALSE,
          height = 500,
          searchable = TRUE,
          defaultColDef = colDef(minWidth = 100, align = "center"),
          columns = list(
            playerName = colDef(minWidth = 250, align = "left")
          ))
```