---
title: AFLW Extended Stats
banner: img/banners/AFLW Logo.png
author: ''
date: '2022-01-10'
slug: aflw-extended-stats
categories: []
tags:
  - AFLW
---

```{r setup, include=FALSE}
library(tidyverse); library(reactable)
```

The AFLW season has kicked off for 2022 and people are starting to get interested in AFLW Fantasy. Moreira's Magic have released a great product that everyone should be getting behind [aflwfantasy.com.au](https://aflwfantasy.com.au/).

We don't envisage providing data tools at this stage but to do our little part in building this community we've collated the advanced stats for Round 1 of the AFLW season and made them downloadable [here](/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv).

The stats available include ruck contests attended, centre bounce attendances (CBAs), kickins taken and played on from, etc.

Head back over after the completion of each round and get updated data (if we don't get too lazy).

* [Download Round 1 Stats](/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv)

```{r, echo=FALSE, message=FALSE}
stats <- read.csv("../../static/data/blog - AFLW Extended Stats Round 1/Extended ALFW Stats Round 1.csv")
stats <- stats %>% filter(fantasy > 0)
reactable(stats,
          pagination = FALSE,
          compact = TRUE,
          highlight = TRUE,
          wrap = FALSE,
          height = 500,
          searchable = TRUE,
          defaultColDef = colDef(minWidth = 100, align = "center"),
          columns = list(
            playerName = colDef(minWidth = 250, align = "left")
          ))
```